---
title: "02_trial_merge.Qmd"
format: html
---


```{r}
library(tidyverse)
library(janitor)
library(here)
library(lme4)
library(lmerTest)
```

Let's read in trial data from the different labs. 

Labs right now are: Senegal, Uganda, Malawi, Rwanda, Ghana, Kenya, and South Africa.

# Ghana

Ghana is in long form. 

```{r}
ghana <- readxl::read_xlsx(here("processed_data", "trials_cleaned", "Omane - Ghana.xlsx")) |>
  janitor::clean_names() |>
  mutate(looking_time_s = as.numeric(lookin_time_s), 
         lab = "Omane - Ghana") |>
  rename(order = test_order) |>
  select(lab, subid, order, trial_type, stimulus, trial_num, looking_time_s, 
         trial_error, trial_error_type) 
```

Just for fun. 

```{r}
ggplot(ghana, 
       aes(x = trial_num, y = looking_time_s, col = trial_type)) + 
  geom_jitter(width = .2, height = 0, alpha = .5) + 
  geom_smooth()
```

```{r}
ggplot(ghana, aes(x = looking_time_s, fill = trial_type)) + geom_histogram(binwidth = 1)
```

# Uganda

Uganda is parsed in `parse_uganda.Qmd`. 

```{r}
uganda <- read_csv(here("processed_data","trials_cleaned","Kizito - Uganda parsed.csv"))
```

```{r}
ggplot(uganda, 
       aes(x = trial_num, y = looking_time_s, col = trial_type)) + 
  geom_jitter(width = .2, height = 0, alpha = .5) + 
  geom_smooth()
```

```{r}
ggplot(uganda, aes(x = looking_time_s, fill = trial_type)) + geom_histogram(binwidth = 1)
```

# Senegal, Malawi, Rwanda, Kenya, Nigeria, South Africa





Next is a set in event format. 

```{r}

# AC: need to fix one file's subid
read_csv(here("processed_data", "trials_cleaned",  "Mushimiyimana - Rwanda", "MBG_IDS_MBG-18_2023-06-11-1105 - feb nzaramba.csv")) %>% 
  mutate(SubjectID = "MBG-18") %>% 
  write_csv(here("processed_data", "trials_cleaned",  "Mushimiyimana - Rwanda", "MBG_IDS_MBG-18_2023-06-11-1105 - feb nzaramba.csv"))

log_labs <- c("Diop - Senegal", "Lamba - Malawi", 
              "Mushimiyimana - Rwanda", "Ziedler - Kenya", "Bentu - Nigeria",
              "Ndhambi - South Africa")




files <- lapply(log_labs, function (x) dir(path = here("processed_data", "trials_cleaned", x)))

log_labs_data_raw <- map_df(1:6, \(x) {
  print(files[x])
  read_csv(here("processed_data", "trials_cleaned", log_labs[x], files[x])) |>
    mutate(lab = log_labs[x])
  }) |>
  filter(!(SubjectID %in% c("Phase", "MBG_IDS"))) |>
  janitor::clean_names()
```
Goal is; select(lab, subid, order, trial_type, stimulus, trial_num, looking_time_s, 
         trial_error, trial_error_type) 

```{r}
log_labs_data <- log_labs_data_raw |>
  filter(lab != "Ndhambi - South Africa") |>
  filter(end_type != "AGAbort") |>
  mutate(subid = subject_id, 
         order = as.numeric(str_sub(order_randomization, 7, 7)), 
         stimulus = str_replace(str_to_lower(stim_name), "\\_", ""), 
         trial_type = case_when(str_detect(stim_name, "IDS") ~ "IDS", 
                                str_detect(stim_name, "ADS") ~ "ADS", 
                                TRUE ~ "training"), 
         trial = as.numeric(trial), 
         trial_num = ifelse(trial < 3, trial - 3, trial - 2), 
         looking_time_s = total_look / 1000, 
         looking_time_diff = (trial_end - trial_start) / 1000, 
         total_center = total_center / 1000, 
         enabled_diff = (look_disabled - look_enabled) / 1000,
         trial_error = NA, 
         trial_error_type = NA) |>
  select(lab, subid, order, trial_type, stimulus, trial_num, looking_time_s, looking_time_diff, total_center, enabled_diff,
         trial_error, trial_error_type) 
```

Now add the trial-order corrected SA data. 

```{r}
log_labs_data_sa <- log_labs_data_raw |>
  filter(lab == "Ndhambi - South Africa") |>
  filter(end_type != "AGAbort") |>
  mutate(subid = subject_id, 
         order = as.numeric(str_sub(order_randomization, 7, 7)), 
         stimulus = str_replace(str_to_lower(stim_name), "\\_", ""), 
         trial_type = case_when(str_detect(stim_name, "IDS") ~ "IDS", 
                                str_detect(stim_name, "ADS") ~ "ADS", 
                                TRUE ~ "training"), 
         trial_num = as.numeric(trial), 
         looking_time_s = total_look / 1000, 
         looking_time_diff = (trial_end - trial_start) / 1000, 
         total_center = total_center / 1000, 
         enabled_diff = (look_disabled - look_enabled) / 1000,
         trial_error = NA, 
         trial_error_type = NA) |>
  select(lab, subid, order, trial_type, stimulus, trial_num, looking_time_s, looking_time_diff, total_center, enabled_diff,
         trial_error, trial_error_type) |>
  group_by(subid)

log_labs_data <- log_labs_data |>
  bind_rows(log_labs_data_sa)
  

```

# Joint analysis

```{r}
d <- bind_rows(ghana, uganda, log_labs_data)
```

We investigate what the different columns mean. Our hypothesis:
- `looking_time_s`: total look = looking time, not including lookaways
- `enabled_diff`: look_disabled - look_enabled = looking time, including lookaways but not attn getters
- `looking_time_diff`: trial_end - trial_start = including lookaways AND attention getters


```{r}
ggplot(d, 
       aes(x = trial_num, y = looking_time_s, col = trial_type)) + 
  geom_jitter(width = .2, height = 0, alpha = .5) + 
  geom_smooth(method = "lm") + 
  facet_wrap(~lab)

```


We have issues with three labs. 

(1) Senegal has a lot of zero looking times. It appears that this is due plausibly to two intersecting issues:
1. a bug in habit such that if you hold down the key forever, you get LT = 0.
2. probable misuse of the software by holding down the key a lot. 

(2) Nigeria has no looking times and their `enabled_diff` is always 18s. I worry that they basically didn't press any keys and just let the experiment run. I also think that they maybe did not use the orders that were assigned. (This is a separate problem). 

(3) South Africa somehow scrambled the trial order so that there are training trials everywhere... I think we can salvage these by just removing the training trials. 


## Senegal

Can we rescue Senegal data? Let's look at their data. They have a few looking times that are not 0 or 18. 

```{r}
ggplot(filter(d, lab == "Diop - Senegal"),
       aes(x = trial_num, y = looking_time_s, col = trial_type)) + 
  geom_jitter(width = .2, height = 0, alpha = .5) + 
  geom_smooth(method = "lm") + 
  facet_wrap(~lab)
```
Let's look at the trial lengths (`enabled_diff`). 

```{r}
ggplot(filter(d, lab == "Diop - Senegal"),
       aes(x = trial_num, y = enabled_diff, col = trial_type)) + 
  geom_jitter(width = .2, height = 0, alpha = .5) + 
  geom_smooth(method = "lm") + 
  facet_wrap(~lab)

```
Reluctantly, I think we can't include these data. 

## South Africa

```{r}
ggplot(filter(d, lab == "Ndhambi - South Africa"),
       aes(x = trial_num, y = looking_time_s, col = trial_type)) + 
  geom_jitter(width = .2, height = 0, alpha = .5) + 
  geom_smooth(method = "lm") + 
  facet_wrap(~lab)
```
This just looks like an order issue. I think we are fine including these if we remove the training trials and report the error. 

# Data passing minimal quality checks


## Extracting excluding trials 

```{r}

d %>% write_csv(here("processed_data/cached_data/unfiltered.csv"))

d_trial_excluded <- d |>
  # this will take out all the ghana trials that marked as having trial error
  filter(!(lab %in% c("Diop - Senegal", "Bentu - Nigeria")),
         looking_time_s <= 0  & trial_type != "training")



d_kids_excluded <- d |>
  filter(!(lab %in% c("Diop - Senegal", "Bentu - Nigeria")),
         looking_time_s > 0,
                   trial_type != "training") %>% 
   mutate(
    
    clean_col = str_replace_all(stimulus, "_", ""),  
    clean_col  = case_when(
      stimulus == "ISD1" ~ "IDS1", 
      TRUE ~ clean_col
      ),
    clean_col = str_to_upper(clean_col),
    stimulus_num = str_extract(clean_col, "\\d+"),
    stimulus_type = str_extract(clean_col, "^[A-Z]+")
  ) %>% 
  distinct(subid, lab, stimulus_num, stimulus_type) %>%
  group_by(subid, lab, stimulus_num) %>%
  summarize(n_types = n(), .groups = "drop") %>%
  # Keep only those stimulus_num where both IDS and ADS are present
  filter(n_types == 2) %>%
  # Now count how many such valid pairs each child has
  count(subid, lab, name = "n_valid_pairs") %>%
  # Optionally filter for children with at least one valid pair
  filter(n_valid_pairs < 1)
  

d_trial_excluded %>% write_csv(here("processed_data/cached_data/excluded_trials.csv"))

d_kids_excluded %>% write_csv(here("processed_data/cached_data/excluded_kids.csv"))
  
  
```


```{r}
d_clean <- d |>
  filter(!(lab %in% c("Diop - Senegal", "Bentu - Nigeria")),
         looking_time_s > 0,
                   trial_type != "training")
  
```

```{r}
ggplot(d_clean, 
       aes(x = trial_num, y = looking_time_s, col = trial_type)) + 
  geom_jitter(width = .2, height = 0, alpha = .5) + 
  scale_y_log10() + 
  facet_wrap(~lab) + 
  geom_smooth(method = "lm") + 
  ylab("Looking Time (log s)") +
  xlab("Trial Number") + 
  ggthemes::theme_few() + 
  ggthemes::scale_color_solarized() +
  theme(legend.position = "bottom")
```
```{r}
write_csv(d_clean, here("processed_data", "trials_cleaned.csv"))
```


# Models

```{r}
d_clean$trial_num_centered <- d_clean$trial_num - 8.5

mod <- lmer(log(looking_time_s) ~ trial_num_centered * trial_type  + 
              (1  | subid) + 
              (trial_type | lab), 
     data = filter(d_clean))
summary(mod)
```


# Data amounts

```{r}
n_trials <- d_clean |>
  group_by(lab, subid) |>
  count() 

ggplot(n_trials, aes(x = n, fill = lab)) + 
  geom_histogram()
```
```{r}
extra_trials <- filter(n_trials, n > 20) |>
  pull(subid)

d_clean |>
  filter(subid %in% extra_trials) 
```

```{r}
d_clean |>
  group_by(lab, subid, trial_num) |>
  count() |>
  filter(n > 1)
```


These come up empty. 

Here's documentation of these changes. 

* Malawi P026 changed to P027 in name and file
* Rwanda deleted MBG1-1 as a duplicate
* Rwanda changed MBG1-9 to MBG1-10 in the file as the name indicated 
* Rwanda changed MBG1-11 to MBG1-11 (from 10) in the file as the name indicated 
* Rwanda changed MBG1-12, 13, and 14 to the correct numbers (was incorrect in file)
* Kenya first instance of MBKENYK009 changed to 008 in both name and file
* Kenya second instance of MBKENYK033 changed to 033B in both name and file
* South africa changed second instance of RSAMB001 to RSAMB001B in name and file
* South africa changed second instance of RSAMB032 to RSAMB0032B in name and file
* South africa changed third instance of RSAMB032 to RSAMB033 in name and file
* South africa changed second instance of RSAMB027 to RSAMB028 in name and file



# Post participant cleaning

AC fixing issues discovered during the fixing participant merge


```{r}
d_clean %>% 
  mutate(subid = case_when(
    subid == "MBG-225" ~ "MBG-25", 
    subid == "RSAMBO14" ~ "RSAMB014", 
    subid == "RSAMBO15" ~ "RSAMB015", 
    subid == "RSAMBO16" ~ "RSAMB016", 
    subid == "RSAMBO17" ~ "RSAMB017", 
    subid == "RSAMBO18" ~ "RSAMB018", 
    subid == "RSAMBO19" ~ "RSAMB019", 
    subid == "RSAMBO20" ~ "RSAMB020",  
   subid == "RSAMBO20" ~ "RSAMB020",  
    subid == "RSAMBO21" ~ "RSAMB021",  
    subid == "RSAMBO22" ~ "RSAMB022",  
    subid == "RSAMBO27" ~ "RSAMB027",  
    subid == "RSAMBOO4" ~ "RSAMB004",  
    subid == "RSAMBOO6" ~ "RSAMB006",  
    subid == "RSASMB028" ~ "RSAMB028",  
    subid == "RSA030" ~ "RSAMB030",
    subid == "RSAMB0004" ~ "RSAMB004",
   subid == "RSAMBO005" ~ "RSAMB005",
    subid == "RSAMB0" ~ "RSAMB008",
   
   subid == "MKENYK0021" ~ "MBKENYK0021",
   
   subid == "IDS_0040" ~ "IDS_040",
   
   subid == "MGB-22" ~ "MBG-22",
   subid == "MGB1-8" ~ "MBG-8",
   
   # make sure naming is consistent from kenya
   subid == "KENNYK_0001" ~ "MBKENYK0001",
   subid == "MBK0018" ~ "MBKENYK0018",
   subid == "MBKENYK_0002" ~ "MBKENYK0002",
   
   TRUE ~ subid
  )) %>% 
  
  mutate(subid = case_when(
      # align the subid in Ghana to be consistent with each other and the ones in the trials data 
    lab == "Omane - Ghana" ~  sub("^mba_(\\d+)$", "mba_p\\1", subid), 
    # Align the Rwanda thing 
    lab == "Mushimiyimana - Rwanda" ~ sub("^MBG1-", "MBG-", subid),
    # Malawi cleanign: 
    lab == "Lamba - Malawi" ~ ifelse(is.na(subid),NA,
    ifelse(
      grepl("_ZOMBA$", subid),
      subid,
      paste0(sub("(_ZOMB)?$", "", subid), "_ZOMBA")
    )
  ),
    
    
    TRUE ~ subid
  )) %>% 
  
  # missing participant information 
  filter(!subid %in% c("RSAMB001B", "RSAMB032B")) %>% 
  
  # needs to be excluded, noted in the participant data file
  filter(!subid %in% c("MBKENYK0001","MBKENYK0018","MBKENYK0002","MBKENYK0003","MBKENYK0004")) %>% 
  
  write_csv(here("processed_data", "trials_cleaned.csv"))
```


AC note: in cleaning the participant data, discovered that the formats of many SA is wrong (O14 instead of 014. change it in the merged file): 

- changed MBG-225 to MBG-25 in trial merge

- O to 0 
"RSAMBO14"  
"RSAMBO15"
"RSAMBO16"  
"RSAMBO17"  
"RSAMBO18"  
"RSAMBO19" 
"RSAMBO20"  
"RSAMBO21"  
"RSAMBO22"  
"RSAMBO27"  
"RSAMBOO4"  
"RSAMBOO6" 
"RSASMB028"

- RSA030 -> RSAMB030
- RSAMB0004 -> RSAMB004
- RSAMBO005 -> RSAMB005
- RSAMB0 -> RSAMB008 (checked the raw file and made sure that the numbers are correct :P)

- dropping RSAMB001B / RSAMB032B because no correspondence in participant data 


- "MKENYK0021"  -> "MBKENYK0021"


