---
title: "02_trial_merge.Qmd"
format: html
---


```{r}
library(tidyverse)
library(janitor)
library(here)
```

Let's read in trial data from the different labs. 

Labs right now are: Senegal, Uganda, Malawi, Rwanda, Ghana, and Kenya. 

# Ghana

Ghana is in long form. First is the easiest one I think. 

```{r}
ghana <- readxl::read_xlsx(here("processed_data", "trials_cleaned", "Omane - Ghana.xlsx")) |>
  clean_names() |>
  mutate(looking_time_s = as.numeric(lookin_time_s)) |>
  rename(order = test_order) |>
  select(lab, subid, order, trial_type, stimulus, trial_num, looking_time_s, 
         trial_error, trial_error_type) 
```

Just for fun. 

```{r}
ggplot(ghana, 
       aes(x = trial_num, y = looking_time_s, col = trial_type)) + 
  geom_jitter(width = .2, height = 0, alpha = .5) + 
  geom_smooth()
```

```{r}
ggplot(ghana, aes(x = looking_time_s, fill = trial_type)) + geom_histogram(binwidth = 1)
```


# Senegal, Malawi, Rwanda, Kenya

Next is a set in event format. 

```{r}
log_labs <- c("Diop - Senegal", "Lamba - Malawi", "Mushimiyimana - Rwanda", "Ziedler - Kenya")

files <- lapply(log_labs, function (x) dir(path = here("processed_data", "trials_cleaned", x)))

log_labs_data_raw <- map_df(1:4, function(x) {
  read_csv(here("processed_data", "trials_cleaned", log_labs[x], files[x])) |>
    mutate(lab = log_labs[x])
  }) |>
  filter(!(SubjectID %in% c("Phase", "MBG_IDS"))) |>
  janitor::clean_names()
```
Goal is; select(lab, subid, order, trial_type, stimulus, trial_num, looking_time_s, 
         trial_error, trial_error_type) 

```{r}
log_labs_data <- log_labs_data_raw |>
  filter(end_type != "AGAbort") |>
  mutate(subid = subject_id, 
         order = as.numeric(str_sub(order_randomization, 7, 7)), 
         stimulus = str_replace(str_to_lower(stim_name), "\\_", ""), 
         trial_type = case_when(str_detect(stim_name, "IDS") ~ "IDS", 
                                str_detect(stim_name, "ADS") ~ "ADS", 
                                TRUE ~ "training"), 
         trial = as.numeric(trial), 
         trial_num = ifelse(trial < 3, trial - 3, trial - 2), 
         looking_time_s = total_look / 1000, 
         looking_time_diff = (trial_end - trial_start) / 1000, 
         total_center = total_center / 1000, 
         enabled_diff = (look_disabled - look_enabled) / 1000,
         trial_error = NA, 
         trial_error_type = NA) |>
  select(lab, subid, order, trial_type, stimulus, trial_num, looking_time_s, looking_time_diff, total_center, enabled_diff,
         trial_error, trial_error_type) 
```


```{r}
d <- bind_rows(ghana, log_labs_data)
```

We investigate what the different columns mean. Our hypothesis:
- total look = looking time, not including lookaways
- look_disabled - look_enabled = looking time, including lookaways but not attn getters
- trial_end - trial_start = including lookaways AND attention getters

The basic issue is that senegal has a lot of zero looking times. It appears that this is due plausibly to two intersecting issues:
1. a bug in habit such that if you hold down the key forever, you get LT = 0.
2. probable misuse of the software by holding down the key a lot. 

```{r}
ggplot(d, 
       aes(x = trial_num, y = looking_time_s, col = trial_type)) + 
  geom_jitter(width = .2, height = 0, alpha = .5) + 
  geom_smooth() + 
  facet_wrap(~lab)

ggplot(d, 
       aes(x = trial_num, y = enabled_diff, col = trial_type)) + 
  geom_jitter(width = .2, height = 0, alpha = .5) + 
  geom_smooth() + 
  facet_wrap(~lab)

```
```{r}
ggplot(d, 
       aes(x = trial_num, y = looking_time_s, col = trial_type)) + 
  geom_jitter(width = .2, height = 0, alpha = .5) + 
  geom_smooth() + 
  facet_wrap(~lab)
```

```{r}
filter(d, lab != "Diop - Senegal", 
       looking_time_s > 2) |>
ggplot(aes(x = trial_num, y = looking_time_s, col = trial_type)) + 
  geom_jitter(width = .2, height = 0, alpha = .5) + 
  scale_y_log10() + 
  geom_smooth(method = "lm")
```

```{r}
mod <- lmer(log(looking_time_s) ~ trial_num * trial_type  + 
              (1 | subid) + 
              (trial_type | lab), 
     data = filter(d, lab != "Diop - Senegal", 
                   looking_time_s > 2, 
                   trial_type != "training"))
summary(mod)
```

